Using Overlapping Communities and Network Structure for Identifying Reduced Groups of Stress Responsive Genes

### 2020-12-07 camila/report/abstract.tex
bmc version

\begin{abstract}
  Abiotic stresses are a main cause of extensive agricultural
  production losses worldwide.  This paper proposes a workflow to
  identify stress responsive genes in organisms: on input RNA
  sequencing read counts measured for genotypes under control and
  treatment conditions, and biological replicates, it outputs a
  collection of characterized genes, potentially relevant to
  treatment. Technically, the proposed approach is both a
  generalization and an extension of WGCNA. It is showcased with a
  systematic study on rice (\textit{Oryza sativa}), a major food
  source that is known to be highly sensitive to salt stress.  A total
  of 6 modules are detected as relevant in the response to salt stress
  in rice: 3 modules of 3 genes each, all associated with shoot K
  content, 2 modules of 3 genes associated with shoot biomass, and 1
  module of 4 genes associated with root biomass. These genes may act
  as potential targets for the improvement of salinity tolerance in
  rice cultivars.
  %% \keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}

>>>>

\begin{abstract}
  This paper proposes a workflow to identify genes responding to a
  specific treatment in an organism, such as abiotic stresses, a main
  cause of extensive agricultural production losses worldwide.  On
  input RNA sequencing read counts (measured for genotypes under
  control and treatment conditions) and biological replicates, it
  outputs a collection of characterized genes, potentially relevant to
  treatment. Technically, the proposed approach is both a
  generalization and an extension of WGCNA; its main goal is to
  identify specific modules in a network of genes after a sequence of
  normalization and filtering steps. In this work, module detection is
  achieved by using Hierarchical Link Clustering, which can recognize
  overlapping communities and thus have more biological meaning given
  the overlapping regulatory domains of systems that generate
  co-expression. Additional steps and information are also added to
  the workflow, where some networks in the intermediate steps are
  forced to be scale-free and LASSO regression is employed to select
  the most significant modules of phenotypical responses to stress.
  %
  %% Biologists may use these genes to develop new crop cultivars with
  %% higher resistance to a specific treatment.
  %
  Finally, the workflow is showcased with a systematic study on rice
  (\textit{Oryza sativa}), a major food source that is known to be
  highly sensitive to salt stress: a total of 6 modules are detected
  as relevant in the response to salt stress in rice;
  %
  %% : 3 modules of 3 genes each, all associated with shoot K content,
  %% 2 modules of 3 genes associated with shoot biomass, and 1 module
  %% of 4 genes associated with root biomass.
  %
  these genes may act as potential targets for the improvement of
  salinity tolerance in rice cultivars.
  %
  The proposed workflow has the potential to ultimately reduce the
  search-space for candidate genes responding to a specific treatment,
  which can considerably optimize the effort, time, and money invested
  by researchers in the experimental validation of stress responsive
  genes.
  \keywords{stress responsive genes \and co-expression network 
            \and overlapping communities \and phenotypic traits
            \and LASSO regression.}
\end{abstract}

### 2020-12-07 camila/report/intro.tex
bmc version

\section{Introduction}
\label{sec.intro}

Abiotic stresses are key factors that can negatively influence plant
development and productivity. They are a main cause of extensive
agricultural production losses worldwide~\cite{todo}. Soil salinity is
one of the most devastating abiotic stresses, causing reduction in the
cultivable land, crop quality, and productivity. It has been estimated
that 20\% of total cultivated and 33\% of irrigated agricultural lands
worldwide are already affected by high salinity. Moreover, due to the
human activities and natural causes, salinized areas are gradually
increasing every year and are expected to reach 50\% by the end of
year 2050~\cite{shrivastava2015soil}. Salinity tolerance and
susceptibility in plants is known to be the result of elaborated
interactions between morphological, physiological, and biochemical
processes that are regulated, in the end, by multiple genes in
different parts of a genome~\cite{reddy2017salt}. Therefore,
identifying groups of stress responsive genes may lead to crop
improvement in terms of salinity tolerance and, ultimately, contribute
solutions to the general problem of food sustainability in the years
to come.

This paper proposes a workflow to identify stress responsive genes in
organisms, which is known to be a complex quantitative trait. It takes
as input RNA sequencing read counts measured for genotypes under
control and treatment conditions (and representing gene expression
profiles of the target organism), and biological replicates.  In order
to discover key genes and their interaction with phenotypes related to
treatment tolerance, the approach requires a collection of phenotypic
traits (under control and under treatment), measured for the given
genotypes. The output of the workflow is a collection of characterized
genes, potentially relevant to treatment, yielding insight on the
possible behavior of specific genes and the role they may play in
functional pathways in response to the studied treatment of the
organism of interest. The proposed workflow can thus take advantage of
transcriptomic data for different organisms and conditions, based on
the current availability of high-throughput technologies that include
microarrays and RNA sequencing, to study the reaction of organisms
under different environmental stimuli, such as salt stress.

Technically, the proposed approach is both a generalization and an
extension of Weighted Gene Co-expression Network Analysis
(WGCNA)~\cite{todo2}, a widely applied workflow that has been
successfully used for identifying target genes related to diseases and
cancer in several organisms~\cite{tian2018identifying}. The general
idea behind each approach is to identify specific modules in a network
of genes after a sequence of normalization and filtering steps. The
proposed approach is considered a \textit{generalization} of WGCNA
because module detection can now recognize overlapping communities,
which may have more biological meaning given the overlapping
regulatory domains of systems that generate
co-expression~\cite{gaiteri2014beyond}. This is achieved by using
Hierarchical Link Clustering (HLC)~\cite{ahn2010link}. It is also an
\textit{extension} of WGCNA because additional steps and information
are added to the workflow: namely, some networks in the intermediate
steps are forced to be scale-free~\cite{todo3} and LASSO
regression~\cite{tibshirani1996regression} is employed to select the
most significant modules of phenotypical responses to stress.  The
advantage of using HLC as clustering method is its ability to detect
overlapping modules, since biological components are involved in
multiple functions and therefore biological communities tend to be
highly overlapping. On the other hand, LASSO is a regularized
regression technique widely used in variable selection, thanks to its
ability to obtain zero regression coefficients for the less relevant
variables~\cite{desboulets2018review}. Moreover, LASSO is especially
useful in problems where the number of variables is much larger than
the number of samples, which may be the case more often than desired.
The proposed workflow is also modular, since other module detection
and selection techniques could be used, instead HLC and LASSO,
respectively.

The proposed workflow is showcased with a systematic study on rice
(\textit{Oryza sativa}), a major food source that is known to be
highly sensitive to salt stress~\cite{chang2019morphological}. RNA-seq
data was accessed from the GEO database~\cite{GEOAcces90:online}
(accession number GSE98455). It corresponds to $57845$ gene expression
profiles of shoot tissues measured for both control and salt condition
in $92$ accessions of the Rice Diversity Panel 1. As output, 6 modules
are detected as relevant in the response to salt stress in rice: 3
modules of 3 genes each, all associated with shoot K content, 2
modules of 3 genes associated with shoot biomass, and 1 module of 4
genes associated with root biomass. These genes may act as potential
targets for the improvement of salinity tolerance in rice
cultivars. From the 19 genes, all but 3 genes (associated with $K$
content), were also identified as deferentially expressed for at least
one of the 92 accessions, suggesting that those genes are strong
candidates as stress responsive genes. Only 2 of the 16 diferentially
expressed genes, both from the module related with shoot biomass, are
named and have an associated protein product: Spermidine
hydroxycinnamoyltransferase 2 (SHT2) and Lipoxygenase. In other words,
further studies are needed to elucidate the detailed biological
function of the remaining 14 genes that have not been named so far,
which may have a potential relevance in stress responsive mechanisms
to salt conditions in rice. The goal is that the results reported in
this paper may allow biologist to develop new rice cultivars with
higher resistance to salinity.

>>>>

\section*{Introduction}
\label{sec.intro}

Stresses are key factors that influence plant
development, often associated to extensive
losses in agricultural production~\cite{mesterhazy2020losses, shrivastava2015soil}. 
Soil salinity is one of the
most devastating abiotic stresses. According to~\cite{shrivastava2015soil}, 
soil salinity contributes to a significant reduction in areas of cultivable
land and crop quality. The study estimates that 
20\% of the total cultivated land worldwide and 
33\% of the total irrigated agricultural land
is affected by high salinity. 
By the end of 2050 areas of high salinity are
expected to reach 50\% of the cultivated land~\cite{shrivastava2015soil}. 
\vspace{0.5cm}

Salinity tolerance and
susceptibility are the result of elaborated
interactions between morphological, physiological, and biochemical
processes, which are regulated by multiple genes in
various parts of the plant genome~\cite{reddy2017salt}. Consequently,
identifying groups of responsive genes is an important step in efforts to
improve crop varieties in terms of salinity tolerance.
This paper proposes a workflow to identify stress responsive genes
associated with a complex quantitative trait.
\vspace{0.5cm}

To discover which genes are associated with a phenotypic response to
treatment the workflow takes as input
the gene expression profiles of the target organism, specifically, 
the RNA sequencing read counts (measured under control
and treatment conditions)
of at least two biological replicates per genotype. It also receives phenotypic data, specifically, observable traits, measured for each genotype under the two conditions. The output of the
workflow is a set of genes which are characterized as potentially relevant
to treatment. 
\vspace{0.5cm}

Broadly speaking, the workflow provides a framework that yields insight into the possible behavior of specific
genes and the role they play in functional pathways in response to
treatment. 
It takes advantage of
the current availability of high-throughput technologies, which
enables the access to transcriptomic data of organisms under different conditions, 
and a better understanding of their reaction under different
environmental stimuli.
\vspace{0.5cm}

The proposed approach is both a generalization and an
extension of the widely applied workflow 
for identifying target genes called 
Weighted Gene Co-expression Network Analysis
(WGCNA)~\cite{langfelder2008wgcna, tian2018identifying}.
Like WGCNA, the general idea behind the proposed
approach is to identify, after a sequence of normalization
and filtering steps, specific modules of overlapping communities
underlying the co-expression network of genes. 
The proposed approach is considered a \textit{generalization}
of WGCNA because module detection recognizes overlapping communities
using Hierarchical Link Clustering (HLC)~\cite{ahn2010link} algorithm.
Conceptually, the generalization takes into account
the overlapping nature of the regulatory domains of the
systems that generate the co-expression network~\cite{gaiteri2014beyond}.
More specifically, overlapping modules allow for 
scenarios where biological components  are involved in
multiple functions.
\vspace{0.5cm}

The workflow is also an \textit{extension} of WGCNA
because two additional constraints are considered:
networks in the intermediate steps are forced to be
scale-free~\cite{barabasi2003scale}, and LASSO
regression~\cite{tibshirani1996regression} selects the
most relevant modules of responsive genes.
The regularized regression technique of LASSO
forces the less relevant variables to be associated to regression
coefficients of value zero~\cite{desboulets2018review}, 
which is of interest
in scenarios where the number of variables is much larger than
the number of samples. This condition is satisfied when the target variables represent the
overlapping communities (obtained with HLC) and the samples represent genotype data,
which is usually a small set due to the high cost of the RNA sequencing process. 
Finally, the proposed workflow is also modular, since other module detection
and selection techniques could be explored instead HLC and LASSO.
\vspace{0.5cm}

The approach is showcased with a systematic study on rice
(\textit{Oryza sativa}), a food source that is known to be
highly sensitive to salt stress~\cite{chang2019morphological}. RNA-seq
data was accessed from the GEO database~\cite{clough2016gene}
(accession number GSE98455). It represents $57845$ gene expression
profiles of shoot tissues measured under control and stress conditions
in $92$ accessions of the Rice Diversity Panel 1~\cite{eizenga2014registration}. A total of 6 modules
are detected as relevant in the response to salt stress in rice: 3
modules, each grouping together 3 genes, are associated to shoot K content; 2
modules of 3 genes, are associated to shoot biomass; and 1 module of 4
genes is associated to root biomass. These genes are potential
targets for experimental validation of salinity tolerance. 
From the 19 genes, 16
are also identified as deferentially expressed for at least
one of the 92 accessions, which re-enforces the labelling of the genes as
stress responsive. Only 2 of the 16 differentially
expressed genes, associated to shoot biomass, are
named and known to produce protein products: spermidine
hydroxycinnamoyltransferase~2 (SHT2) and lipoxygenase. 
Further studies are needed to elucidate the detailed biological
functions of the remaining 14 genes and their role
in the mechanisms that respond 
to salt conditions.

\paragraph{Paper Outline.}
%The remainder of the paper is organized as
%follows. Section~\ref{sec.prelim} gathers preliminaries on gene
%co-expression networks, HLC, and LASSO. The proposed workflow is
%presented in Section~\ref{sec.framework}, with especial focus on the
%logical steps of the process and the internal structures supporting
%the approach. Section~\ref{sec.case} presents the case study on the
%identification of rice genes that are sensitive to salt
%stress. Section~\ref{sec.concl} draws some conclusions and future research directions.

The remainder of the paper is organized as
follows. The \hyperref[sec.prelim]{Preliminaries} section gathers foundations on gene
co-expression networks, HLC, and LASSO. The proposed workflow is
presented in \hyperref[sec.framework]{the Workflow} section, which enphasizes on the
logical steps of the data analysis process and the internal structures supporting
the approach. The \hyperref[sec.case]{Case Study} section presents an application of the workflow for the
identification of rice genes that are sensitive to salt
stress. Finally, the \hyperref[sec.concl]{Concluding Remarks} section draws some conclusions and future research directions.

### 2020-12-07 camila/report/main.tex
bmc version

\begin{document}
%
\title{Using Complex Networks for Identifying Potentially New Saline
  Stress Responsive Genes in Rice}
%
\author{Camila Riccio \and Jorge Finke \and Camilo Rocha}
%% \author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
%% Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
%% Third Author\inst{3}\orcidID{2222--3333-4444-5555}}

@@

\institute{Pontificia Universidad Javeriana, Cali, Colombia
  %% \email{lncs@springer.com}
}  
%
\maketitle

>>>>

\begin{document}
%
\title{Using Overlapping Communities and Network Structure for
  Identifying Reduced Groups of Stress Responsive Genes}
% \titlerunning{ }
\titlerunning{Using Network Features for Identifying Stress Responsive Genes}
%
%% \title{Using complex networks for identifying a reduced group of stress responsive genes}
%
\author{Camila Riccio \and Jorge Finke \and Camilo Rocha}
%% \author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
%% Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
%% Third Author\inst{3}\orcidID{2222--3333-4444-5555}}

@@

\institute{Pontificia Universidad Javeriana, Cali, Colombia
  \email{\{camila.riccio,jfinke,camilo.rocha\}@javerianacali.edu.co}
}  
%
\maketitle

### 2020-12-07 camila/report/prelim.tex
bmc version

\section{Preliminaries}
\label{sec.prelim}

This section presents preliminaries on networks, hierarchical link
clustering and the LASSO linear regression technique.

\subsection{Co-expression network}

A network is an undirected graph $G=(V,E)$ where
${V=\{v_1,v_2,\ldots,v_{n}\}}$ is a set of \textit{vertices} or
\textit{nodes} and ${E=\{e_1,e_2,\ldots,e_q\}}$ is a set of
\textit{edges} or \textit{links} that connect vertices. In a gene
co-expression network, each node corresponds to a gene. A pair of
genes is connected if they show similar expression
patterns. A simple and unweighted network can be represented by an
adjacency matrix $A \in \{0,1\}^{n \times n}$ that is symmetric with a
positive one in the positions $(v_i,v_j)$ and $(v_j,v_i)$ whenever
there is an edge connecting vertices $v_i$ and $v_j$, and zeros
elsewhere. Co-expression networks are of biological interest because
the co-expressed genes are usually controlled by the same
transcriptional regulatory pathway, functionally related or members of
the same pathway or metabolic complex.

\subsection{Hierarchical Link Clustering}

The Hierarchical Link Clustering (HLC) algorithm was proposed by Ahn
et al.~\cite{ahn2010link}. The HLC approach represents communities as
groups of links (rather than nodes), and each node inherits all
memberships of its links and can thus belong to multiple, overlapping
communities. It maps links to nodes and connects them if a pair of
links shares a node. The similarity between two links $e_{ik}$ and
$e_{jk}$ is computed using the Jaccard index
%
\begin{equation}\label{eq:jaccard}
  S(e_{ik},e_{jk}) = \frac{\vert n(i) \cap n(j) \vert}{\vert n(i) \cup n(j) \vert},
\end{equation}
%
where $n(i)$ denotes the set containing exactly node $i$ and its
neighbors. The algorithm uses single-linkage hierarchical clustering
to build a dendrogram in which each leaf is a link from the original
network and branches represent link communities. Hierarchical
clustering algorithms repeatedly merge groups until all elements are
members of a single cluster. For the purpose of finding meaningful
communities, it is crucial to know where to partition the
dendrogram. In this case, the most relevant communities are
established at the maximal partition density $D$, a function based on
link density inside communities measuring the quality of a link
partition. The partition density $D$ has a single global maximum along
the dendrogram in almost all cases, because its value is the average
density at the top of the dendrogram (a single giant community with
every link and node) and it is very small at the bottom of the
dendrogram (most communities consists of a single link). In
particular, it is the case that $D = 1$ when every community is a
fully connected clique and $D = 0$ when each community is a tree. If a
community is less dense than a tree (i.e. when the community subgraph
has disconnected components), then such a community contribute
negatively to $D$, which can take negative values. The minimum density
inside a community is $-2/3$, given by one community of two
disconnected edges. Since $D$ is the average of the intra-community
density, there is a lower bound of $-2/3$ for $D$. Computing $D$ at
each level of the link dendrogram can help the purpose of picking the
best level to cut, although meaningful structure could exist above or
below the threshold. The output of cutting is a set of node clusters,
where each node can participate in multiple communities.

\subsection{Least Absolute Shrinkage Selector Operator}

The Least Absolute Shrinkage Selector Operator (LASSO) is a
regularized linear regression technique. It combines a regression
model with a procedure of contraction of some parameters towards zero
and selection of variables, imposing a restriction or a penalty on the
regression coefficients. In other words, LASSO solves the least
squares problem with restriction on the $ L_1$-norm of the coefficient
vector. It can be especially useful to solve problems where the number
of variables (e.g., genes) $ n $ is much greater than the number of
samples $ m $ (i.e., $ n \gg m $).

Consider a dataset consisting of $m$ samples, each of which consists
of $n$ covariates and a single outcome. Let $y_i$ be the outcome and
$x_i := (x_1,...,x_n)$ be the covariate vector for the $i$-th
sample. The objective of LASSO is to solve
%
\begin{equation}
\min \left\lbrace\sum_{i=1}^{m}{\left( y_i-\sum_{j=1}^n{\beta_j
    x_{ij}}\right)^2} \right\rbrace \quad , \quad \textrm{subject to}
\quad \sum_{j=1}^n\abs{\beta_j}\leq s.
\end{equation}

Equivalently, in the Lagrangian form, it minimizes

\begin{equation}
  \sum_{i=1}^{p}{\left( y_i-\sum_{j=1}^n{\beta_j x_{ij}}\right)^2} +
  \lambda \sum_{j=1}^n\abs{\beta_j}
\end{equation}
%
where $s$ is the regularization penalty and $\lambda \geq 0$ is the
corresponding Lagrange multiplier. Since the $\lambda$ value
determines the degree of penalty, the accuracy of the model depends on
its choice. Cross-validation is often used to select the
regularization parameter, choosing the one that minimizes the
mean-squared error.

>>>>

\section*{Preliminaries}
\label{sec.prelim}

This section presents preliminaries on networks, the clustering algorithm HLC,
and the linear regression technique LASSO.

\subsection*{Co-expression network}

Consider a network as an undirected graph $G=(V,E)$ where
${V=\{v_1,v_2,\ldots,v_{n}\}}$ is a set of $n$ \textit{vertices} (or
\textit{nodes}) and ${E=\{e_1,e_2,\ldots,e_q\}}$ is a set of $q$
\textit{edges} (or \textit{links}) that connect vertices. In a
co-expression network of genes, each node corresponds to a gene, and
a link indicates similar expression pattern between two genes.
The network can be represented by an adjacency matrix $A
\in \{0,1\}^{n \times n}$ that is symmetric. A matrix entry in 
positions $(v_i,v_j)$ and $(v_j,v_i)$ is equal to $1$ whenever there is an edge
connecting vertices $v_i$ and $v_j$, and equal to $0$
otherwise. Co-expression networks are of biological interest because
adjacent nodes in the network represent co-expressed genes that are
usually controlled by the same transcriptional regulatory pathway,
functionally related, or members of the same pathway or metabolic
complex~\cite{FIONDA2019915}.

\subsection*{Hierarchical Link Clustering}

The Hierarchical Link Clustering (HLC) algorithm partitions
groups of links (rather than nodes), each node inherits all
memberships of its links and can belong to multiple, overlapping
communities~\cite{ahn2010link}. More specifically, HLC evaluates the similarity 
between links if they share a particular node.
Consider a pair of links $e_{ik}$ and $e_{jk}$, which are adjacent to node $k$.
The similarity between $e_{ik}$ and $e_{jk}$ is defined based on the Jaccard index as
%
\begin{equation}\label{eq:jaccard}
  S(e_{ik},e_{jk}) = \frac{\vert \ \eta(i) \cap \eta(j) \ \vert}{\vert \ \eta(i) \cup \eta(j) \ \vert},
\end{equation}
%
where $\eta(i)$ denotes the set containing node $i$ and its
neighbors. The algorithm uses single-linkage hierarchical clustering
to build a dendrogram in which each leaf is a link from the
network, and branches represent link communities. 
%Hierarchical
%clustering algorithm repeatedly merge groups until all elements are
%members of a single cluster.
\vspace{0.5cm}

The threshold where to cut the dendogram is defined
based on the average density of links inside communities (partition density).
For $G=(V,E)$ and a partition of the links into $c$ subsets, the partition density is computed as

\begin{equation}
D = \frac{2}{\vert E \vert} \sum_c \vert E_c \vert \frac{\vert E_c \vert - \vert V_c \vert + 1 }{(\vert V_c \vert -1)(\vert V_c \vert -2)}
\end{equation}

Note that in most cases, the partition density $D$ has a single
global maximum along the dendrogram. 
If the dendogram is cut at the top, then $D$ represents the
average link density of a single
giant community.  If the dendogram is cut at the
bottom, then most communities consists of a single
link. In other words, note that $D = 1$ when every community
is a clique and $D = 0$ when each community is a
tree. If a community is less dense than a tree (i.e., when the
community subgraph has disconnected components), then such a community
contributes negatively to $D$, which can take negative values. The
minimum density inside a community is $-2/3$, given by one community
of two disconnected edges. Since $D$ is the average of the
intra-community density, there is a lower bound of $-2/3$ for
$D$. By computing $D$ at each level of the dendrogram, 
the level that maximizes partition density can be found 
(nonetheless meaningful
structure could exist above or below the threshold).

% figure 0
\begin{figure}[htbp]
  \centering
    \includegraphics[clip,width=0.96\textwidth]{figures/figure0.png}
  \caption[Example of a full link dendrogram (left) and partition density (right)]%
  {Example of a full link dendrogram (left) and partition density (right), borrowed from~\cite{ahn2010link}.}
  \label{fig:hlc_density}
\end{figure}

The output of
cutting is a set of node clusters, where each node can participate in
multiple communities.

\subsection*{Least Absolute Shrinkage Selector Operator (LASSO)}

LASSO is a
regularized linear regression technique. By combining a regression
model with a procedure of contraction of some parameters towards $0$,
LASSO imposes a restriction (or a penalty) on
regression coefficients. In other words, LASSO solves the least
squares problem with restriction on the $ L_1$-norm of the coefficient
vector. In particular, the approach is especially useful in scenarios where the number
of variables $ c $ is much greater than the number of
samples $ m $ (i.e., $ c \gg m $).
\vspace{0.5cm}

Consider a dataset of $m$ samples, consisting each
of $c$ covariates and a single outcome. Let $y_i$ be the outcome and
$x_i := (x_{i1},...,x_{ic})$ be the covariate vector for the $i$-th
sample. The objective of LASSO is to solve
%
\begin{equation}
\min \left\lbrace\sum_{i=1}^{m}{\left( y_i-\sum_{j=1}^c{\beta_j
    x_{ij}}\right)^2} \right\rbrace \quad , \quad \textrm{subject to}
\quad \sum_{j=1}^c\abs{\beta_j}\leq s.
\end{equation}

where $s$ is the regularization penalty. 
Equivalently, in the Lagrangian form, LASSO minimizes

\begin{equation}
  \sum_{i=1}^{p}{\left( y_i-\sum_{j=1}^c{\beta_j x_{ij}}\right)^2} +
  \lambda \sum_{j=1}^c\abs{\beta_j}
\end{equation}
%
where $\lambda \geq 0$ is the
corresponding Lagrange multiplier. Since the value of the regularization
parameter $\lambda$
determines the degree of penalty and the accuracy of the model,
cross-validation is used to select the
regularization parameter that minimizes the
mean-squared error.
